---
title: "Challenge 7"
author: "Emily Lo"
format:
  html:
    theme: cerulean
    self-contained: true
    code-overflow: wrap
    code-tools: true
    code-fold: true
execute: 
  echo: true
  errors: true
  warning: false
  messages: false
---

Reading the data

```{r}
library(tidyverse)
library(here)
library(ggtext)
```

```{r}
fishes_data <- read_csv(here("week 7","lab 7","BlackfootFish.csv"))
```

Part 1 : Use your function to modify the `length` and `weight` columns of the `BlackfootFish` dataset, removing values you believe are "unreasonable.

```{r}
#used this code to find unreasonable values
fishes_data |> ggplot(aes(x=weight)) + 
  geom_histogram()

```

```{r}

impossible_fixer <- function(data, min, max) {
  clean_data <- case_when(
    data < min ~ NA_real_,
    data > max ~ NA_real_,
    TRUE ~ data)
  return(clean_data)}

fishes_data <- fishes_data |> 
  mutate(length = impossible_fixer(data = length,
                                        min = 100, 
                                        max = 500),
         weight = impossible_fixer(data= weight,
                                        min = 100, 
                                        max = 1000))

```

```{r}
fishes_data <- fishes_data |>
  mutate(length_mm = length * 25.4, 
         weight_grams = weight * 453.592)
  
```

Part 2: Write a function which calculates the condition index of a fish, given inputs of weight and length.

```{r}
fish_condition_index <- function(weight, length) {
return(weight/length^3 * 100)}
  
fishes_data <- fishes_data |>
  mutate(condition_index = fish_condition_index(weight_grams, 
                                                length_mm))

```

Part 3: Make a thoughtful visualization of how fish conditions have varied over the duration of this study.

```{r}
fish_visual <- fishes_data |> 
  group_by(year, 
           species, 
           trip) |> 
  summarize(average_condition_index = mean(condition_index, 
                                           na.rm = TRUE))

```

```{r}
my_palette <- c("#00FFFF", "#90EE90")

ggplot(data = fish_visual, 
       aes(x = year, 
           y = average_condition_index,
           color = as.factor(trip))) + 
  geom_line() + 
  facet_wrap(~ species, 
             ncol = 2) + 
  labs(y = "", 
       x = "Year") +
  scale_color_manual(values = my_palette) +
    labs(title = str_c("Trout Condition Index of ",
      "<span style = 'color : #00FFFF'>**Trip 1**</span>",
      " and <span style = 'color:#90EE90'>**Trip 2**</span>"),
      subtitle = "By Year") +
  theme_gray() +
  theme(panel.grid = element_blank(),
        axis.line = element_line(size = 0.5, 
                                 color = "pink"),
        plot.title = element_markdown(size = 15), 
        axis.text.y = element_text(size = 12), 
        plot.background = element_rect(fill = "grey90", color = "black", linewidth = 1), 
        strip.background = element_rect(colour = "red", 
                                        fill = "pink"),
        legend.position = "none")
```
