---
title: "Lab 3"
author: "Emily Lo"
format:
  html:
    theme: cerulean
    self-contained: true
    code-overflow: wrap
    code-tools: true
    code-fold: true
execute: 
  echo: true
  errors: true
  warning: false
  messages: false
---

# Getting Started

[Download starter .qmd file](lab3-hiphop-lyrics.qmd)

::: callout-caution
# Add your preferred YAML

The downloaded file **does not** contain an YAML, you need to include your own. This allows you to specify how you would like your rendered HTML to look! The sky is the limit!
:::

[Download `hiphop.csv`](hiphop.csv)

::: callout-warning
Save **both** your .qmd file and your data set in the same folder within your Stat 331 directory! **DO NOT** open your .qmd file straight from your downloads folder.
:::

## Tips for Formatting your Lab

::: {.callout-tip collapse="true"}
-   The first chunk of your Quarto document should be to *declare your libraries* (probably only `tidyverse` for now).
-   The second chunk of your Quarto document should be to *load in your data* (using the `here()` function!).
-   Make sure you address **all the questions** in these instructions.
-   If a question requires **any** type of calculation, **you should provide code for your answer!**
-   I have provided hints about functions that might be useful to you. You are **not required** to use these functions.
-   You may have to Google to solve some of these!
-   Be sure to **save** your work regularly (`Ctrl/Cmd + S` or `File > Save` or the floppy disk icon)
-   Be sure to **render** your file every so often, to check for errors and make sure it looks nice.
    -   Make your Quarto document does not contain `View(dataset)` or `install.packages("package")`, both of these will prevent rendering.
    -   Check your Quarto document for moments when you looked at the data by typing the name of the data frame. Leaving these in means the whole dataset will print out and this looks unprofessional.
    -   If you are unable to finish due to errors in your code, remember that you can still submit an HTML file if you specify `error: true` in your `execute` options (in the YAML).
:::

# Data Set

The data set `hiphop` contains results from a study conducted by a linguist at the University of Minnesota. The researcher was interested in predicting musical taste based on familiarity with African American English (AAE). 168 subjects participated in the study, and each was asked to define 64 different AAE terms. The definitions given were used to create a `familiarity` score for each subject for each term. This score quantifies how well the subject knew the term on a scale of 1-5 (1 = not at all, 5 = very well). Before tackling the problems, study the description of each variable [here](http://conservancy.umn.edu/bitstream/handle/11299/116327/5/explanationAAEHiphopChesley.txt).

**1. Load the appropriate R packages and import the data set, `hiphop`.**

```{r packages}
# code chunk for loading packages and importing the data
library(tidyverse) 
library(here)
hiphop <- read_csv(here("week 3","lab3", "hiphop.csv"))
```

## Summary

**2. Provide a brief overview (2-4 sentences) of the data set.**

The data set consists of 10752 rows and 38 columns. The data consists of a variety of types of variables such as characters, double class numeric values, and dates.

```{r dataset-explore}
# you may want to use code to answer this question
dim(hiphop)
class(hiphop)
view(hiphop)
```

::: callout-note
It is always good practice to start an analysis by getting a feel for the data and providing a quick summary for readers. You do not need to show any source code for this question, although you probably want to use code to get some information about the data set.
:::

**3. What are the rows of this data set?**

**The rows of this data set represent a different word for each subject. For example, the subject p1 has one row for the "ghost" word and another row for the "5-0" word.**

::: callout-warning
It is not one person per row!
:::

```{r rows}
# you may want to use code to answer this question
view(hiphop)
```

## Cleaning the Data

**4. Missing values for some of the variables were replaced with other values. How were missing values replaced? What do you believe are some benefits and drawbacks of doing this?**

**Missing values were replaced with mean values for the column. A benefit is that it was easy to compute and is somewhat accurate. However, this data point can be very far and can skew the data.**

**5. Clean the data set in whichever ways you see fit. This might mean adjusting *variable type*, for example from `character` to `factor`, or dealing with missing data. Assign your cleaned data set to a new data set named `hiphop_clean` -- use this data set going forward.**

::: callout-tip
Helpful functions: `mutate()`, `as.factor()`

Likert scales (1-5) will be read by R as numeric values and MUST be converted to factors.

It would be most efficient to use `across()` in combination with `mutate()` to complete this task.
:::

```{r}
# code chunk for Q5
hiphop_clean <- hiphop |>
  mutate(word = as.factor(word),
         sex = as.factor(sex),
         ethnic = as.factor(ethnic),
         across(where(is.numeric), as.factor)) |>
  drop_na(word, sex, ethnic)
```

# Data Summaries

**6. How many unique AAE words were studied in this data set?**

64 unique AAE words were studied in this data set.

::: callout-tip
Helpful functions: `distinct()`, `count()`
:::

```{r}
# code chunk for Q6
distinct(hiphop_clean, word) |>
count()
```

**7. Make a new variable that re-categorizes `ethnic` into only two groups, "white" and "non-white", to simplify your data.**

::: callout-tip
Helpful functions: `mutate()`, `if_else()`
:::

```{r}
# code chunk for Q7

hiphop_clean <- hiphop_clean |> 
  mutate(ethnicity = if_else(ethnic == "white", "white", "non-white"))

```

**8. It is fairly common for researchers to collapse ethnic or racial categories similar to what you just did. What are some issues with representing the data in this way?**

White vs. non-white does not give enough detailed information about the research to the common people. It is also important and will also give researchers sufficient information when they know what exact ethnicities are included in the "non-white" ethnicity.

**9. What are the demographics of the people in this study? Investigate the variables `sex`, `age`, and `ethnic` and summarize your findings in 1-3 complete sentences.**

```{r}
# code chunk for Q9
demographics <- hiphop_clean |>
  select(sex, age, ethnic) |>
  distinct(.keep_all = TRUE) 
```

Age of people range from 16 to 42 years old. Ethnicity is majority white over non-whites. There are more females than males.

**10. Make at least two plots to display the demographic information of the subjects in this study.**

::: callout-note
You do not need to discuss these plots, but make sure they are appropriate to the data types and have informative titles and axis labels. Feel free to use the skills you learned in Challenge 2 to enhance your plots!
:::

```{r}
# code chunk for Q10
#plot 1 
ggplot(data = hiphop_clean, mapping = aes(x=sex, y=age, color = ethnic)) +  
        geom_boxplot(alpha=0.2) + facet_wrap(~ethnic) +
        labs(x="Sex", y="Age", color= "", 
        title = "Distribution between age and sex by Ethnicity")
```

```{r}
ggplot(data = hiphop_clean, mapping = aes(x=sex, y=age, fill="ethnic")) + 
  geom_boxplot() + geom_jitter(col="tomato", alpha=0.1) + facet_wrap(~ethnic) +
  labs(x="Sex", y="Age", title = "Distribution of Sex and Age within each Ethnicity")
```

## Familiar words

For each demographic group listed below, determine which word(s) in this study was(were) the most **and** least familiar on average.

::: callout-tip
Helpful functions: `filter()`, `group_by()`, `summarize()`, `slice_max()`, `slice_min()`

Useful variables: `word`, `familiarity`, `sex`, `age`, `ethnic`
:::

**11. People below the age of 20.**

```{r}
# code chunk for Q11
young_people <- hiphop_clean |>
  filter(age < 20)|>
  group_by(word) |>
  summarize(Mean = mean(familiarity)) |>
  slice_max(Mean, n = 1) |>
  group_by(word)
```

**12. Non-white women.**

```{r}
# code chunk for Q12
non_whitewomen <- hiphop_clean |>
  filter(sex == "Female", ethnic != "white")

most_familiar_words <- non_whitewomen |>
  group_by(word) |>
  summarize(mean_familiarity = mean(familiarity)) |>
  slice_max(mean_familiarity, n = 1)

least_familiar_words <- non_whitewomen |>
  group_by(word) |>
  summarize(mean_familiarity = mean(`familiarity`)) |>
  slice_min(mean_familiarity, n = 1)

```

**13. White men above the age of 30.**

```{r}
# code chunk for Q13
white_men_above_30 <- hiphop_clean |>
  filter(sex == "Male", ethnicity == "white", age > 30) 

most_familiar_words <- white_men_above_30 |>
  group_by(word) |>
  summarize(mean_familiarity = mean(familiarity)) |>
  slice_max(mean_familiarity, n = 1)

least_familiar_words <- white_men_above_30 |>
  group_by(word) |>
  summarize(mean_familiarity = mean(familiarity)) |>
  slice_min(mean_familiarity, n = 1)
```

## Study Subjects

A joke among the [Tidy Tuesday](https://www.tidytuesday.com/) community is that Justin Bieber was one of the subjects in this study. Bieber, a white male, from a relatively small town (10,000-60,000 people) in Ontario would have been 17-23 at the time of the study.

**14. Determine which subject you believe is secretly Bieber, justify your answer.**

::: callout-tip
Refer again to the data set description. There is another clue about Bieber's identity.
:::

```{r}
# code chunk for Q14
bieber <- hiphop_clean |> 
  filter(sex == "Male", age >= 17, age <= 23, city >= 10000, city <= 60000, bieber > 4)

bieber
```

# Lab 3 Submission

You will submit **only** your rendered HTML file. Your HTML file is required to have the following specifications in the YAML options (at the top of your document):

-   be self-contained (`self-contained: true`)
-   include your source code (`code-tools: true`)
-   include all your code and output (`echo: true`)
-   include **no** messages from loading in packages or the data (`messages: false`)

**If any of the options are not included, your Lab 3 or Challenge 3 assignment will receive an "Incomplete" and you will be required to submit a revision.**

# Challenge 3: Group Comparisons & Data Ethics

## Published Comparisons

In the published article ([You Know What It Is: Learning Words through Listening to Hip-Hop](https://journals.plos.org/plosone/article/file?id=10.1371/journal.pone.0028248&type=printable)), the author presents a series of comparisons about the variables which most explain an individual's familiarity with African American English (AAE).

Let us instead compare the number of artists participants reported liking for each genre. Specifically, you will determine which music genre most differentiates each of the demographic groups provided.

> Which genre had much higher average (mean or median) reported artists in one group than the other.
>
> -   Male versus Female
> -   White versus Non-White

::: callout-tip
You might find it helpful to first create a new data set with only the variables you are interested in! Look at the Music Variables in the data set description along with the demographics of interest.

Helpful functions: `select()`, `group_by()`, `summarize()`, `across()`

Other useful operations in R: `mean()`, `diff()`, `abs()`, `which.max()`
:::

```{r}
male_vs_female <- hiphop_clean |>
select(sex, intl, vocal, classical, folk, rock, country, pop, alternative, hiphop, unclassifiable)  |> group_by(sex)  |> summarize(mean_intl = mean(intl), mean_vocal = mean(vocal), mean_classical = mean(classical),
mean_folk = mean(folk), mean_rock = mean(rock), mean_country = mean(country), mean_pop = mean(pop), mean_alternative = mean(alternative), mean_hiphop = mean(hiphop),mean_unclassifiable = mean(unclassifiable))

male_vs_female
```

## Study Design Critique -- Data Ethics

Myself, members of the Tidy Tuesday community, and previous 331 students have voiced concerns regarding the design and social context of this study.

You've already read the [data description (link)](http://conservancy.umn.edu/bitstream/handle/11299/116327/5/explanationAAEHiphopChesley.txt) regarding how participants were recruited for participation in this study. **Now**, you need to read additional details regarding aspects of the study in the published paper: [You Know What It Is: Learning Words through Listening to Hip-Hop](https://journals.plos.org/plosone/article/file?id=10.1371/journal.pone.0028248&type=printable).

> Based on the design of this study and its context (African American English), what are **at least two** concerns you have? Keep in mind this critique aligns with conversations regarding data ethics. Thus, your concerns need to address the racial aspects of the design of this study.

# Challenge 3 Submission

Your challenge should be submitted as a **separate file**, **not** at the bottom of your Lab 3 file. Please submit your rendered HTML file. You can copy and paste this code into a **new** Quarto file. Your Challenge 3 submission should only included code necessary for completing the Challenge, nothing else.

You will submit **only** your rendered HTML file. Your HTML file is required to have the following specifications in the YAML options (at the top of your document):

-   be self-contained (`self-contained: true`)
-   include your source code (`code-tools: true`)
-   include all your code and output (`echo: true`)
-   include **no** messages from loading in packages or the data (`messages: false`)

**If any of the options are not included, your Lab 3 or Challenge 3 assignment will receive an "Incomplete" and you will be required to submit a revision.**
