---
title: "practice activity 9.1 regression"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(here)
library(dplyr)
```

# **1 Data**

The dataset [mystery_animal.csv](https://atheobold.github.io/stat-331/9_simulation-regression/Practice%20Activity/mystery_animal.csv) contains the weights of a particular animal before and after a year of eating only roasted duck.

```{r}
mystery_animal <- read_csv(here("week 9", 
                                "practice activity", 
                                "mystery_animal.csv"))
```

# **2 Linear Regression**

Fit a linear regression to determine if the duck diet is associated with the animal gaining weight, losing weight, or neither.

```{r}
mystery_animal_lm <- lm(weight_after ~ weight_before,
                        data = mystery_animal)
```

# **3 Residuals**

Plot the residuals (observed value minus predicted value) of your linear model versus `weight_before`.

```{r}
mystery_animal_lm |>
  broom::augment() |>
  ggplot(mapping = aes(y = .resid, x = .fitted)) +
  geom_point()
```

# **4 Mystery Animal**

What animal do you see?

shark
