---
title: "Practice Activity 2"
author: "Emily Lo"
format:
  html:
    theme: cerulean
    self-contained: true
    code-overflow: wrap
    code-tools: true
    code-fold: true
execute: 
  echo: true
  errors: true
  warning: false
  messages: false
---

```{r}
library(tidyverse)
```

```{message <- read_csv(here::here("week 5","practice activity","scrambled_message.txt"))}
```

Question 1

892 characters are in the scrambled message

```{r}
message %>% 
  mutate(length = str_length(Word)) %>% 
  summarise(total_length = sum(length))

```

Question 2

127 words are in the scrambled message

```{r}
message |> 
  filter(str_detect(Word, pattern = "\\w")) %>% 
  count() 
```

Question 3

Print out every piece of a word in the scrambled message that starts with the letter "m".

```{r}
words <- c(message$Word)
str_subset(words, pattern = "^m")
```

Question 4

Print out every piece of a word in the scrambled message that ends with the letter "z"

```{r}
str_subset(words, pattern = "z$")
```

Question 5

Print out every punctuation symbol in the scrambled message.

```{r}

words %>% str_extract_all(pattern = "[[:punct:]]") %>%
  purrr::map(~ data_frame(punctuation = .)) %>%
  bind_rows()

```

Decode a message

```{r}
words <- c(message$Word)
library(stringr)
words <- str_trim(words)


words <- sapply(words, function(x) substr(x, 1, 16))

words <- gsub("ugh[!.?-@]*", "", words, ignore.case = TRUE)

words <- gsub("aa", "ee", words)

words <- gsub("z", "t", words)

words <- gsub("b$", "y", words)

words <- gsub("^k", "v", words)

message <- str_c(words, collapse = " ")

view(message)
```
